Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-02-06 paper/main.tex
working on methods

The results thus far inform us about which aspects of the dynamic content in the episode participants watched were preserved or altered in participants' memories.  We next carried out a series of analyses aimed at understanding which brain structures might facilitate these preservations and transformations between the external world and memory.  In one analysis, we sought to identify brain structures that were sensitive to the dynamic unfolding of the video's content, as characterized by its topic trajectory.  We used a searchlight procedure to identify clusters of voxels whose activity patterns displayed a proximal temporal correlation structure (as participants watched the video) matching that of the original video's topic proportions (Fig.~\ref{fig:brainz}A; see \textit{Methods} for additional details).  In a second analysis, we sought to identify brain structures whose responses (during video viewing) reflected how each participant would later structure their \textit{recalls} of the video.  We used an analogous searchlight procedure to identify clusters of voxels whose proximal temporal correlation matrices matched that of the topic proportions for each individual's recall (Figs.~\ref{fig:brainz}B; see \textit{Methods} for additional details).  To ensure our searchlight procedure identified regions \textit{specifically} sensitive to the temporal structure of the video or recalls (i.e., rather than those with a temporal autocorrelation length similar to that of the video/recalls), we performed a phase shift-based permutation correction (see \textit{Methods} for additional details).  Specifically, we circularly shifted the timeseries of the topic trajectory by a random number of timepoints, recomputed the shifted trajectory's correlation matrix, and again performed our searchlight analysis on this permuted data.  We then $z$-scored the ``real" searchlight results at each voxel against the null distribution of (100) permuted results.  In Figure~\ref{fig:brainz}, only voxels whose activity pattern reflected the ``real" video/recall timeseries more closely than 95\% of the permuted results are shown.  As shown in Figure~\ref{fig:brainz}C, the video-driven searchlight analysis revealed a distributed network of regions including \textbf{\textit{????}}, suggesting that these regions may play a role in processing information relevant to the narrative structure of the video. As shown in Figure~\ref{fig:brainz}D, the analysis revealed a network of regions including \textbf{\textit{????}}, suggesting that these regions facilitate a person-specific transformation of one's experience into memory.  In identifying regions whose responses to ongoing experiences reflect how those experiences will be remembered later, this latter analysis extends classic \textit{subsequent memory analyses}~\citep[e.g.,][]{PallWagn02} to domain of naturalistic stimuli.
% ------ NOTE: ------
% Potentially need to fill in ????s with names of structures? But maybe not; better to leave without inference because of neurosynth analysis?
% - Paxton

@@

Our work casts remembering as reproducing (behaviorally and neurally) the topic trajectory, or shape, of an experience.  This view draws inspiration from prior work aimed at elucidating the neural and behavioral underpinnings of how we process dynamic naturalistic experiences and remember them later.  One approach to identifying neural responses to naturalistic stimuli (including experiences) entails building a model of the stimulus and searching for brain regions whose responses are consistent with the model.  In prior work, a series of studies from Uri Hasson's group~\citep{LernEtal11, SimoEtal16, ChenEtal17, BaldEtal17, ZadbEtal17} have extended this approach with a clever twist: rather than building an explicit stimulus model, these studies instead search for brain responses (while experiencing the stimulus) that are reliably similar across individuals.  So called \textit{inter-subject correlation} (ISC) and \textit{inter-subject functional connectivity} (ISFC) analyses effectively treat other people's brain responses to the stimulus as a ``model'' of how its features change over time.  By contrast, in our present work we used topic models to construct an explicit content model directly from the stimulus (i.e., the topic trajectory of the video).  Projecting each participant's recall into a space shared by both the stimulus and other participants then allows us to compare recalls both directly to the stimulus and to each other.  Similarly, prior work introducing the use of HMMs to discover latent event structure in naturalistic stimuli and recall \citep{BaldEtal17} used a between-subjects cross-validation approach to identify event boundaries shared across participants, and between stimulus and recall.  Our framework allows us to break from the restriction of a common, shared event-timeseries and identify the unique \textit{resolution} of each participant's recall event structure, and how that may differ from the video and each other.  
% @JEREMY @ANDY PLEASE CHECK THIS TO MAKE SURE I'M NOT MISCHARACTERIZING BALDETAL17

@@

\subsection*{Modeling the dynamic content of the video and recall transcripts}
\subsubsection*{Topic modeling}
The input to the topic model we trained to characterize the dynamic content of the video comprised 998 hand-generated annotations of short (mean: 2.96s) scenes spanning the video clip~(\citealp{ChenEtal17} generated 1000 annotations total; we removed two referring to the break between the first and second scan sessions, during which no fMRI data was collected).  The features annotated included: narrative details (a sentence or two describing what happened in that scene); whether the scene took place indoors or outdoors; names of any characters that appeared in the scene; name(s) of characters in camera focus; name(s) of characters who were speaking in the scene; the location (in the story) that the scene took place; camera angle (close up, medium, long, top, tracking, over the shoulder, etc.); whether music was playing in the scene or not; and a transcription of any on-screen text.  We concatenated the text for all of these features within each segment, creating a ``bag of words'' describing each scene.  We then re-organized the text descriptions into overlapping sliding windows of 50 scenes each.  In other words, we created a ``context" for each scene comprising the text descriptions of the preceding 25 scenes, the present scene, and the following 24 scenes.  To model the ``context" at the beginning and end of the video (i.e., within 25 scenes of the beginning or end), we created overlapping sliding windows that grew in size from one scene to the full length, then similarly tapered their length at the end.  This bore the additional benefit of representing each scene's description in the text corpus an equal number of times.
% ------ NOTE: ------
% need to refine explanation here a little bit
% - Paxton

We trained our model using these overlapping text samples with \texttt{scikit-learn}~\citep[version 0.19.1; ][]{PedrEtal11}, called from our high-dimensional visualization and text analysis software, \texttt{HyperTools}~\citep{HeusEtal18a}.  Specifically, we used the \texttt{CountVectorizer} class to transform the text from each window into a vector of word counts (using the union of all words across all scenes as the ``vocabulary,'' excluding English stop words); this yielded a number-of-windows by number-of-words \textit{word count} matrix.  We then used the \texttt{LatentDirichletAllocation} class (topics=100, method=`batch') to fit a topic model~\citep{BleiEtal03} to the word count matrix, yielding a number-of-windows (1047) by number-of-topics (100) \textit{topic proportions} matrix.  The topic proportions matrix describes which mix of topics (latent themes) is present in and around each scene.  Next, we transformed the topic proportions matrix to match the 1976 fMRI volume acquisition times.  We assigned each topic vector to the timepoint midway between the beginning of the first scene and the end of the last scene in its corresponding sliding text window.  We then transformed these timepoints to units of TRs and interpolated the dynamic topic proportions matrix to obtain number-of-TRs (1976) by number-of-topics (100) matrix.

We created similar topic proportions matrices using hand-annotated transcripts of each participant's recall of the video~\citep[annotated by ][]{ChenEtal17}.  We tokenized the transcript into a list of sentences, and then re-organized the list into overlapping sliding windows spanning 10 sentences each (and analogously tapered the lengths of the first and last 10 sliding windows).  In turn, we transformed each window's sentences into a word count vector (using the same vocabulary as for the video model).  We then used the topic model already trained on the video scenes to compute the most probable topic proportions for each sliding window.  This yielded a number-of-windows (range: 83--312) by number-of-topics (100) topic proportions matrix for each participant.  These reflected the dynamic content of each participant's recalls.  Note: for details on how we selected the video and recall window lengths and number of topics, see \textit{Supporting Information} and Figure~\topicopt.

@@

We used an optimization procedure to select the appropriate $K$ for each topic proportions matrix.  Prior studies on narrative structure and processing have shown that we both perceive and internally represent the world around us at multiple, hierarchical timescales (e.g., \citealp{HassEtal08, LernEtal11, HassEtal15, ChenEtal17, BaldEtal17, BaldEtal18}).  However, for the purposes of our framework, we sought to identify the single timescale of event-representations that is emphasized \textit{most heavily} in the temporal structure of the video and each participant's recalls.  We quantified this as the set of $K$ event boundaries that yielded the maximal distinctiveness between the content (i.e., topics) within each event and that in all other events. Specifically, we computed (for each matrix)
\[
  \argmax_K \left[W_{1}(a, b)\right],
\]
where $a$ was the distribution of correlations between the topic vectors of timepoints within the same state and $b$ was the average correlation between the topic vectors of timepoints within \textit{different} states.  For each possible $K$, we computed the first Wasserstein distance ($W_{1}$; also known as ``earth mover's distance"; \citealp{Dobr70, RamdEtal17}) between these distributions, and chose the $K$-value that yielded the greatest difference.  Figure~\ref{fig:model}B displays the event boundaries returned for the video, and Figure~\corrmats~displays the event boundaries returned for each participant's recalls (See Fig.~\kopt~for the optimization functions for the video and recalls).  After obtaining these event boundaries, we created stable estimates of each topic proportions matrix by averaging the topic vectors within each event.  This yielded a number-of-events by number-of-topics matrix for the video and recalls from each participant.

>>>>

The results thus far inform us about which aspects of the dynamic content in the episode participants watched were preserved or altered in participants' memories.  We next carried out a series of analyses aimed at understanding which brain structures might facilitate these preservations and transformations between the external world and memory.  In one analysis, we sought to identify brain structures that were sensitive to the dynamic unfolding of the video's content, as characterized by its topic trajectory.  We used a searchlight procedure to identify clusters of voxels whose activity patterns displayed a proximal temporal correlation structure (as participants watched the video) matching that of the original video's topic proportions (Fig.~\ref{fig:brainz}A; see \textit{Methods} for additional details).  In a second analysis, we sought to identify brain structures whose responses (during video viewing) reflected how each participant would later structure their \textit{recalls} of the video.  We used an analogous searchlight procedure to identify clusters of voxels whose proximal temporal correlation matrices matched that of the topic proportions for each individual's recall (Figs.~\ref{fig:brainz}B; see \textit{Methods} for additional details).  To ensure our searchlight procedure identified regions \textit{specifically} sensitive to the temporal structure of the video or recalls (i.e., rather than those with a temporal autocorrelation length similar to that of the video/recalls), we performed a phase shift-based permutation correction (see \textit{Methods} for additional details).  Specifically, we circularly shifted the timeseries of the topic trajectory by a random number of timepoints, recomputed the shifted trajectory's correlation matrix, and again performed our searchlight analysis on this permuted data.  We then $z$-scored the ``real" searchlight results at each voxel against the null distribution of (100) permuted results.  In Figure~\ref{fig:brainz}, only voxels whose activity pattern reflected the ``real" video/recall timeseries more closely than 95\% of the permuted results are shown.  As shown in Figure~\ref{fig:brainz}C, the video-driven searchlight analysis revealed a distributed network of regions that may play a role in processing information relevant to the narrative structure of the video.  Similarly, the recall-driven searchlight analysis revealed a second diffuse network of regions (Fig.~\ref{fig:brainz}D) that may facilitate a person-specific transformation of one's experience into memory.  In identifying regions whose responses to ongoing experiences reflect how those experiences will be remembered later, this latter analysis extends classic \textit{subsequent memory analyses}~\citep[e.g.,][]{PallWagn02} to domain of naturalistic stimuli.

@@

Our work casts remembering as reproducing (behaviorally and neurally) the topic trajectory, or shape, of an experience.  This view draws inspiration from prior work aimed at elucidating the neural and behavioral underpinnings of how we process dynamic naturalistic experiences and remember them later.  One approach to identifying neural responses to naturalistic stimuli (including experiences) entails building a model of the stimulus and searching for brain regions whose responses are consistent with the model.  In prior work, a series of studies from Uri Hasson's group~\citep{LernEtal11, SimoEtal16, ChenEtal17, BaldEtal17, ZadbEtal17} have extended this approach with a clever twist: rather than building an explicit stimulus model, these studies instead search for brain responses (while experiencing the stimulus) that are reliably similar across individuals.  So called \textit{inter-subject correlation} (ISC) and \textit{inter-subject functional connectivity} (ISFC) analyses effectively treat other people's brain responses to the stimulus as a ``model'' of how its features change over time.  By contrast, in our present work, we use topic models to construct an explicit content model directly from the stimulus (i.e., the topic trajectory of the video).  Projecting each participant's recall into a space shared by both the stimulus and other participants then allows us to compare recalls both directly to the stimulus and to each other.  Similarly, prior work introducing the use of HMMs to discover latent event structure in naturalistic stimuli and recall \citep{BaldEtal17} used between-subjects cross-validation to identify event boundaries shared across participants, and between stimulus and recall.  Our framework allows us to break from the restriction of a common, shared event-timeseries and identify the unique \textit{resolution} of each participant's recall event structure, and how that may differ from the video and that of other participants.  
% @JEREMY @ANDY PLEASE CHECK THIS TO MAKE SURE I'M NOT MISCHARACTERIZING BALDETAL17

@@

\subsection*{Modeling the dynamic content of the video and recall transcripts}
\subsubsection*{Topic modeling}
The input to the topic model we trained to characterize the dynamic content of the video comprised 998 hand-generated annotations of short (mean: 2.96s) scenes spanning the video clip~(\citealp{ChenEtal17} generated 1000 annotations total; we removed two referring to the break between the first and second scan sessions, during which no fMRI data was collected).  We concatenated the text for all of the annotated features within each segment, creating a ``bag of words'' describing each scene and performed some minor preprocessing (e.g., stemming possessive nouns and removing punctuation).  We then re-organized the text descriptions into overlapping sliding windows spanning (up to) 50 scenes each.  In other words, we created a ``context" for each scene comprising the text descriptions of the preceding 25 scenes, the present scene, and the following 24 scenes.  To model the ``context" for scenes near the beginning and end of the video (i.e., within 25 scenes of the beginning or end), we created overlapping sliding windows that grew in size from one scene to the full length, then similarly tapered their length at the end.  This additionally ensured that each scene's content was represented in the text corpus an equal number of times.
% In other words, one sample contained the text from scenes 1-50, the next contained the text from scenes 2-51, and so on.  In order to ensure each scene's content was represented in these 
% ------ NOTE: ------
% need to refine explanation here a little bit
% - Paxton

We trained our model using these overlapping text samples with \texttt{scikit-learn}~\citep[version 0.19.1; ][]{PedrEtal11}, called from our high-dimensional visualization and text analysis software, \texttt{HyperTools}~\citep{HeusEtal18a}.  Specifically, we used the \texttt{CountVectorizer} class to transform the text from each window into a vector of word counts (using the union of all words across all scenes as the ``vocabulary,'' excluding English stop words); this yielded a number-of-windows by number-of-words \textit{word count} matrix.  We then used the \texttt{LatentDirichletAllocation} class (topics=100, method=`batch') to fit a topic model~\citep{BleiEtal03} to the word count matrix, yielding a number-of-windows (1047) by number-of-topics (100) \textit{topic proportions} matrix.  The topic proportions matrix describes the gradually evolving mix of topics (latent themes) present in each scene.  Next, we transformed the topic proportions matrix to match the 1976 fMRI volume acquisition times.  We assigned each topic vector to the timepoint (in seconds) midway between the beginning of the first scene and the end of the last scene in its corresponding sliding text window.  By doing so, we warped the linear temporal distance between consecutive topic vectors to align with the inconsistent temporal distance between consecutive annotations (whose durations varied greatly).  We then rescaled these timepoints to 1.5s TR units, and used linear interpolation to estimate a topic vector for each TR.  This resulted in a number-of-TRs (1976) by number-of-topics (100) matrix.

We created similar topic proportions matrices using hand-annotated transcripts of each participant's recall of the video~\citep[annotated by ][]{ChenEtal17}.  We tokenized the transcript into a list of sentences, and then re-organized the list into overlapping sliding windows spanning (up to) 10 sentences each, analogously to how we parsed the video annotations.  In turn, we transformed each window's sentences into a word count vector (using the same vocabulary as for the video model), then used the topic model already trained on the video scenes to compute the most probable topic proportions for each sliding window.  This yielded a number-of-windows (range: 83--312) by number-of-topics (100) topic proportions matrix for each participant.  These reflected the dynamic content of each participant's recalls.  Note: for details on how we selected the video and recall window lengths and number of topics, see \textit{Supporting Information} and Figure~\topicopt.

@@

We used an optimization procedure to select the appropriate $K$ for each topic proportions matrix.  Prior studies on narrative structure and processing have shown that we both perceive and internally represent the world around us at multiple, hierarchical timescales (e.g., \citealp{HassEtal08, LernEtal11, HassEtal15, ChenEtal17, BaldEtal17, BaldEtal18}).  However, for the purposes of our framework, we sought to identify the single timeseries of event-representations that is emphasized \textit{most heavily} in the temporal structure of the video and of each participant's recall.  We quantified this as the set of $K$ states that maximized the similarity between topic vectors for timepoints comprising each state, while minimizing the similarity between topic vectors for timepoints across different states.  Specifically, we computed (for each matrix)
\[
  \argmax_K \left[W_{1}(a, b)\right],
\]
where $a$ was the distribution of within-state topic vector correlations, and $b$ was the distribution of across-state topic vector correlations .  We computed the first Wasserstein distance ($W_{1}$; also known as ``earth mover's distance"; \citealp{Dobr70, RamdEtal17}) between these distributions for a large range of possible $K$-values (range [2,50]), and selected the $K$ that yielded the greatest difference.  Figure~\ref{fig:model}B displays the event boundaries returned for the video, and Figure~\corrmats~displays the event boundaries returned for each participant's recalls (See Fig.~\kopt~for the optimization functions for the video and recalls).  After obtaining these event boundaries, we created stable estimates of each topic proportions matrix by averaging the topic vectors within each event.  This yielded a number-of-events by number-of-topics matrix for the video and recalls from each participant.
